RNeXML: a package for reading and writing richly annotated phylogenetic, character, and trait data in R

### 2015-04-27 manuscripts/manuscript.Rmd
Adds reference to VTO.

A detailed account of how the NeXML standard addresses these and
other relevant challenges can be found in @Vos_2012. In brief,
NeXML was designed with the following important properties. First,
NeXML is defined by a precise grammar that can be programmatically
**validated**; i.e., it can be verified whether a file precisely
follows this grammar, and therefore whether it can be read (parsed) 
without errors by software that uses the NeXML grammar (e.g. RNeXML)
is predictable. 
Second, NeXML is **extensible**: a user
can define representations of new, previously unanticipated information
(as we will illustrate) without violating its defining grammar. Third
and most importantly, NeXML is rich in **computable semantics**: it is
designed for expressing metadata such that machines can understand
their meaning and make inferences from it. For example, OTUs in a tree
or character matrix for frog species can be linked to concepts in a
formally defined hierarchy of taxonomic concepts such as the
Vertebrate Taxonomy Ontology [@Midford_2014], which enables a machine
to infer that a query for amphibia is to include the frog data in what
is returned. (For a more broader discussion of the value of such
capabilities for evolutionary and biodiversity science we refer the
reader to @Parr2011.)

>>>>

A detailed account of how the NeXML standard addresses these and
other relevant challenges can be found in @Vos_2012. In brief,
NeXML was designed with the following important properties. First,
NeXML is defined by a precise grammar that can be programmatically
**validated**; i.e., it can be verified whether a file precisely
follows this grammar, and therefore whether it can be read (parsed) 
without errors by software that uses the NeXML grammar (e.g. RNeXML)
is predictable. 
Second, NeXML is **extensible**: a user
can define representations of new, previously unanticipated information
(as we will illustrate) without violating its defining grammar. Third
and most importantly, NeXML is rich in **computable semantics**: it is
designed for expressing metadata such that machines can understand
their meaning and make inferences from it. For example, OTUs in a tree
or character matrix for frog species can be linked to concepts in a
formally defined hierarchy of taxonomic concepts such as the
Vertebrate Taxonomy Ontology [@Midford2013], which enables a machine
to infer that a query for amphibia is to include the frog data in what
is returned. (For a more broader discussion of the value of such
capabilities for evolutionary and biodiversity science we refer the
reader to @Parr2011.)
